---
title: "Download GEO annotations"
author: "`r Sys.info()[['user']]`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  github_document:
    toc: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

This notebook takes the [../annotations/geo_accession_numbers.csv] file as input and produces two tables:
1. A table of all samples (uniquely identified by GSM numbers) [../annotations/geo_annotations.tsv]. This uses the `GEOquery` R package.
2. A table of all SRA runs (uniquely identified by SRR numbers) [../annotations/sra_annotations.tsv]. This uses the `reutils` R package.

# Load libraries and define analysis-specific variables
```{r}
library(GEOquery)
library(reutils)
library(xml2)
library(glue)
library(tidyverse)
```

# Load GEO datasets that we want to download annotations and download GEO record
```{r}
geo_accessions <- read_csv("../annotations/geo_accession_numbers.csv", comment = "#") %>% 
  mutate(geo = map(gse, . %>% getGEO(GSEMatrix = FALSE))) %>%
  print()
```

# Define functions for parsing specific meta-data values
```{r}
get_bioproject_number <- function(geo) {
  Meta(geo) %>% 
  enframe("annotation", "value") %>%
  unnest() %>%
  mutate(bioproject_number = str_extract(value, "PRJNA[:digit:]+")) %>%
  filter(!is.na(bioproject_number)) %>%
  pull(bioproject_number)
}

get_gsm_meta <- function(gsm) {
  Meta(gsm) %>% 
  enframe("annotation", "value") %>%
  unnest()
}

get_srx_number <- function(gsm) {
  gsm %>% 
    mutate(srx = str_extract(value, "SRX[:digit:]+")) %>%
    filter(!is.na(srx)) %>%
    pull(srx)
}
    
get_gsm_title <- function(gsm) {
  gsm %>% 
    filter(annotation == "title") %>% 
    pull(value) %>% 
    janitor::make_clean_names()
}

get_gsm_organism <- function(gsm) {
  gsm %>% 
    filter(str_detect(annotation, "organism")) %>% 
    pull(value) %>% 
    janitor::make_clean_names()
}

get_gsm_celltype <- function(gsm) {
  gsm %>%
    mutate(celltype = str_extract(value, "293|((?<=cell line: |tissue: |cell type: ).+)")) %>%
    filter(!is.na(celltype)) %>%
    slice(1) %>% 
    pull(celltype) %>% 
    janitor::make_clean_names()
}
```

# Parse Experiment and Project annotations
```{r}
geo_annotations <- geo_accessions %>% 
  mutate(prjna = map_chr(geo, get_bioproject_number)) %>% 
  mutate(gsm = map(geo, GSMList)) %>% 
  select(-geo) %>% 
  mutate(gsm_number = map(gsm, names)) %>%
  unnest() %>%
  mutate(gsm = map(gsm, get_gsm_meta)) %>%
  mutate(sample_title = map_chr(gsm, get_gsm_title)) %>%
  mutate(srx = map_chr(gsm, get_srx_number)) %>%
  mutate(organism = map_chr(gsm, get_gsm_organism)) %>%
  mutate(cell_type = map_chr(gsm, get_gsm_celltype)) %>%
  select(-gsm) %>%
  rename(gsm = gsm_number) %>%
  select(gsm, gse, prjna, srx, cell_type) %>%
  group_by(gsm) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(cell_type = if_else(str_detect(cell_type, "embryonic_kidney"), "HEK293T", cell_type)) %>% 
  write_tsv("../annotations/geo_annotations.tsv") %>%
  print()
```

# Get SRA Run information for the BioProject number associated with each study
```{r}
esearch_query <- geo_annotations %>% 
  distinct(prjna) %>% 
  # filter(prjna %in% c("PRJNA279465", "PRJNA555528")) %>%
  pull() %>% 
  paste0(collapse = " or ") %>% 
  print()

esearch_result <- esearch(esearch_query, db = "sra")
efetch_result <- efetch(esearch_result, db = "sra")
```

# Parse XML file to extract SRA information
```{r}
sra_info <- read_xml(content(efetch_result, as = "text"))

sra_annotations <- sra_info %>% 
  xml_find_all("//EXPERIMENT") %>% 
  map_df(xml_attrs) %>%
  rename(srx = accession, gsm = alias) %>% 
  mutate(sample_name = map(gsm, function (x) xml_text(xml_find_all(sra_info, glue('//SAMPLE[@alias="{x}"]/TITLE'))))) %>%
  mutate(data = map(srx, function (x) bind_rows(xml_attrs(xml_find_all(sra_info, glue('//EXPERIMENT_REF[@accession="{x}"]/..')))))) %>%
  unnest() %>%
  select(gsm, srx, sample_name, accession, total_spots, total_bases) %>%
  type_convert() %>%
  rename(srr = accession) %>%
  mutate(sample_name = janitor::make_clean_names(sample_name)) %>%
  select(srr, everything()) %>%
  left_join(select(geo_annotations, gsm, gse, cell_type), by = "gsm") %>%
  select(-srx, -total_spots, -total_bases) %>%
  left_join(select(geo_accessions, -geo), by = "gse") %>%
  filter(str_detect(sample_name, tolower(study_subset)) | is.na(study_subset)) %>%
  write_tsv("../annotations/sra_annotations.tsv") %>%
  print()
```
